---
title: "Portfolio 8"
author: "Fei Xie"
date: "5/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Better visualization of the change of CO2 emission and forest land and their relationship

Here we also use CO2 and Forest data from 1990 to 2015 (abstracted from GapMinder) to see the change of CO2 emission and forest land in there years and examine their relationship. 

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(dplyr)
library(rworldmap)
```

```{r load-data, message=FALSE}
CO2 <- read_csv("data/co2_emissions_tonnes_per_person.csv", na = c("",".","NA"))
```

## To which degree the CO2 emission has changed from 1990 to 2015 in each country?

### Absolute Change

```{r add-change}
CO2$change <- (CO2$`2015` - CO2$`1990`)
```


```{r plot-change}
ggplot(CO2, 
       aes(x = change))+
  geom_histogram(color = "blue")


ggplot(CO2, 
       aes(x= CO2$`1990`, 
           y = CO2$`2015`))+
  geom_point(color = "darkred",
             size = 3,
           alpha = CO2$change)

sPDF <- joinCountryData2Map(CO2,
                            #use country name instead of country code to join data
                            joinCode="NAME"
              , nameJoinColumn = "country")

mapCountryData( sPDF
              , nameColumnToPlot="change" ,
              catMethod="pretty",
              colourPalette='heat',addLegend='TRUE',
              mapTitle= "World map about the CO2 emission change from 1990 to 2015",
              missingCountryCol = gray(.8)
              )
```

### Relative Change

```{r plot-change-ratio}
CO2$ratio <- (CO2$`2015` - CO2$`1990`)/CO2$`1990`

ggplot(CO2, 
       aes(x = ratio))+
  geom_histogram(color = "blue")


sPDF <- joinCountryData2Map(CO2,
                            #use country name instead of country code to join data
                            joinCode="NAME"
              , nameJoinColumn = "country")

mapCountryData( sPDF
              , nameColumnToPlot="ratio" ,
              catMethod="pretty",
              colourPalette='heat',addLegend='TRUE',
              mapTitle= "World map about the CO2 emission change ratio from 1990 to 2015",
              missingCountryCol = gray(.8)
              )
```

## CO2 now predicted by CO2 before? Slope for each country
```{r slop-function}
slope <- function(x) {
  if (all (is.na(x)))
    return(NA)
  else
    return(coef(lm(I(1:26)~x))[2])
}
```

```{r calculate-each-country}
CO2$slope <- 
  apply (CO2[, c("1990", "1991","1992",
                 "1993", "1994","1995",
                 "1996", "1997","1998",
                 "1999", "2000","2001",
                 "2002", "2003", "2004",
                 "2005", "2006", "2007",
                 "2008", "2009", "2010",
                 "2011", "2012", "2013",
                 "2014", "2015")],1,slope)
```

```{r plot-slopes}
ggplot(CO2, 
       aes(x= CO2$`1990`, 
           y = CO2$`2015`))+
  geom_point(color = "darkred",
             size = 3,
           alpha = -CO2$slope/10)+
  labs(title = "Can the CO2 emission before predict 2015 CO2 emission in countries",
       x = "CO2 emission in 1990",
       y = "CO2 emission in 2015",
       color = "strength of prediction")
```

```{r worldmap-slopes}
sPDF <- joinCountryData2Map(CO2,
                            #use country name instead of country code to join data
                            joinCode="NAME"
              , nameJoinColumn = "country")

mapCountryData( sPDF
              , nameColumnToPlot="slope" ,
              catMethod="pretty",
              colourPalette='heat',addLegend='TRUE',
              mapTitle= "World map about the CO2 emission 1990 predicting emission 2015",
              missingCountryCol = gray(.8)
              )
```
